---
title: "Figure 8"
output:
  html_document:
    df_print: paged
---
Script to regenerate figure 8, the comparison of MOI inferred by SS-Cat and RM-Cat in 3 populations in Uganda.

Load the data frame needed to generate the figures. 
The data frame includes inferred MOI from SS-Cat and RM-Cat from 128 MCMC runs in 3 populations.

The columns of large_cat represent host id, inferred MOI, inference method, population (there are 2 columns because left_join was used when preparing the data frame), total count of rows in each population by each method, and the weight of each row (1/count, used in plotting fraction).
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)

load("figure8.Rdata")
head(large_cat)
```

Plotting figure 8.
```{r fig.height=6, fig.width=14, message=FALSE, warning=FALSE}
color<-c("SS-Bin"="#D55E00","SS-Pois"="#E69F00","SS-NegBin"="#F0E442","SS-Cat"="#CC79A7","RM-Prop"="#0072B2","RM-Cat"="#009E73")

#Get the median value of inferred MOI by both methods in each population
med_cat <- large_cat %>% group_by(pop.x,method) %>% summarise(med = median(inferred_MOI))
med_cat$method<-factor(med_cat$method,levels=c("SS-Cat","RM-Cat"))

large_cat$method<-factor(large_cat$method,levels=c("SS-Cat","RM-Cat"))

g<-ggplot(large_cat)+geom_bar(aes(weight=weight,x=factor(inferred_MOI),fill=method),position=position_dodge2(preserve="single"),alpha=0.7)+
  theme_bw()+xlab("Inferred MOI")+ylab("Frequency")+theme_clean()+
  scale_x_discrete(labels = ~ ifelse(seq_along(.x) %% 2 == 1, .x, ""))+
  scale_y_continuous(limits=c(0,0.6),breaks=seq(0.1,1,0.1))+theme(plot.title = element_text(hjust = 0.5))+
  geom_segment(data=med_cat,aes(x=med,y=0,xend=med,yend=0.56,color=method),linetype=2,linewidth=1)+scale_fill_manual(values=color)+
  geom_text(data=med_cat,aes(label = med,x=med,y=0.58,color=method),size=8,show.legend = F)+
  theme(plot.title=element_text(size=28,face="bold"), axis.text=element_text(size=24),
        axis.title=element_text(size=28,face="bold"),plot.background = element_blank(),legend.key.size = unit(1.5, "cm"),
        legend.position = c(0.9,0.75),legend.title=element_text(size=24,face="bold"),legend.text=element_text(size=24),
        strip.background =element_rect(fill="grey50"),strip.text = element_text(colour = 'white'))+scale_color_manual(values=color)+ 
  facet_grid(~factor(pop.x,levels=c("Nagongera","Walukuba","Kihihi")),scales="free")+theme(strip.text.x = element_text(size=24,face="bold"),strip.text.y = element_text(size=24,face="bold"))+theme(panel.spacing.x = unit(2, "lines"))
g
```

