---
title: "Figure S1"
output:
  html_document:
    df_print: paged
---

Script to regenerate figure S1, the mean estimation errors for each run (Stability), first averaged over hosts or sites.

Load the data frame needed to generate the figures. 
The data frame moiError includes true and inferred MOI, heterozygosity for each host after merging identical strains, in 128 runs for each method under each scenario. The data frame af includes the corresponding true and inferred minor allele frequency.

The columns of moiError represent task id (1-128), host id, true MOI, inferred MOI, true heterozygosity, inferred heterozygosity, error in MOI (inferred-true), scenario of transmission intensity, inference method, and class of input data (read counts or categorical data).


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(patchwork)
library(cowplot)

load("figure4_S1.Rdata")
head(moiError)
```

The columns of af represent task id (1-128), SNP site id, true minor allele frequency, inferred minor allele frequency, scenario of transmission intensity, inference method, and class of input data (read counts or categorical data).
```{r message=FALSE, warning=FALSE}
head(af)
```

Get means of MOI, heterozygosity and minor allele frequency over hosts or sites for each run.
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#Get means of MOI and heterozygosity over hosts for each run
moiErrorStat<- moiError%>%ungroup()%>%group_by(task_id,method,scenario)%>%
  summarize(moiError = mean(error),hetError = mean(inferred_het-true_het))

# define different method categories
moiErrorStat$methodCat <- ifelse(moiErrorStat$method %in% c("SS-NegBin","SS-Pois","SS-Bin","RM-Prop"),"Read Counts","Categorical")
moiErrorStat$methodCat <- factor(moiErrorStat$methodCat,levels = c("Read Counts","Categorical"))

#Get mean of minor allele frequency over sites for each run
afErrorStat<- af%>%ungroup()%>%group_by(task_id,method,scenario)%>%
  summarize(afError = mean(inferred_freq-true_freq))

# define different method categories
afErrorStat$methodCat <- ifelse(afErrorStat$method %in% c("SS-NegBin","SS-Pois","SS-Bin","RM-Prop"),"Read Counts","Categorical")
afErrorStat$methodCat <- factor(afErrorStat$methodCat,levels = c("Read Counts","Categorical"))
```

Plotting figure S1
```{r fig.height=15, fig.width=12, message=FALSE, warning=FALSE, paged.print=FALSE}
color<-c("SS-NegBin"="#D55E00","SS-Pois"="#E69F00","SS-Bin"="#F0E442","SS-Cat"="#CC79A7","RM-Prop"="#0072B2","RM-Cat"="#009E73")

# MOI subplot
p1<-ggplot(moiErrorStat,aes(methodCat,y=moiError,color=method))+
  geom_hline(yintercept = 0,linewidth=1.5,color="grey")+
  stat_boxplot(geom = "errorbar",position = position_dodge2(preserve = "single")) + 
  geom_boxplot(aes(fill=method),alpha=0.3,position = position_dodge2(preserve = "single"))+
  scale_color_manual(values=color)+
  scale_fill_manual(values=color)+
  facet_grid(~scenario)+
  ylab("mean error of\ninferred MOIs")+
  xlab("")+
  theme_cowplot()+
  guides(color = guide_legend(nrow = 1))+
  theme(legend.position = "top",
        #legend.text = element_text(size=14)
        panel.grid.major.y = element_line(linetype = "dotted",color="grey")
  )+
  theme(plot.tag=element_text(size=28),
        plot.title=element_text(size=18,face="bold"), axis.text=element_text(size=18),
        legend.title=element_text(size=18,face="bold"),legend.text=element_text(size=18),legend.background=element_blank(),
        axis.title=element_text(size=18,face="bold"),plot.background = element_blank(),
        strip.text.x = element_text(size=20,face="bold"),
        strip.background =element_rect(fill="grey50"),strip.text = element_text(colour = 'white'))

# Het subplot
p2<-ggplot(moiErrorStat,aes(methodCat,y=hetError,color=method))+
  geom_hline(yintercept = 0,linewidth=1.5,color="grey")+
  stat_boxplot(geom = "errorbar",position = position_dodge2(preserve = "single")) + 
  geom_boxplot(aes(fill=method),alpha=0.3,position = position_dodge2(preserve = "single"))+
  scale_color_manual(values=color)+
  scale_fill_manual(values=color)+
  facet_grid(~scenario)+
  ylab("mean error of\ninferred heterozygosities")+
  xlab("")+
  theme_cowplot()+
  theme(legend.position = "none",panel.grid.major.y = element_line(linetype = "dotted",color="grey"))+
  theme(plot.tag=element_text(size=28),
        plot.title=element_text(size=18,face="bold"), axis.text=element_text(size=18),
        legend.title=element_text(size=18,face="bold"),legend.text=element_text(size=18),legend.background=element_blank(),
        axis.title=element_text(size=18,face="bold"),plot.background = element_blank(),
        strip.text.x = element_text(size=20,face="bold"),
        strip.background =element_rect(fill="grey50"),strip.text = element_text(colour = 'white'))

# MAF subplot
p3<-ggplot(afErrorStat,aes(methodCat,y=afError,color=method))+
  geom_hline(yintercept = 0,linewidth=1.5,color="grey",lty=1)+
  stat_boxplot(geom = "errorbar",position = position_dodge2(preserve = "single")) + 
  geom_boxplot(aes(fill=method),alpha=0.3,position = position_dodge2(preserve = "single"))+
  scale_color_manual(values=color)+
  scale_fill_manual(values=color)+
  facet_grid(~scenario)+
  ylab("mean error of\ninferred allele frequencies")+
  xlab("")+
  theme_cowplot()+
  theme(legend.position = "none",panel.grid.major.y = element_line(linetype = "dotted",color="grey"))+
  theme(plot.tag=element_text(size=28),
        plot.title=element_text(size=18,face="bold"), axis.text=element_text(size=18),
        legend.title=element_text(size=18,face="bold"),legend.text=element_text(size=18),legend.background=element_blank(),
        axis.title=element_text(size=18,face="bold"),plot.background = element_blank(),
        strip.text.x = element_text(size=20,face="bold"),
        strip.background =element_rect(fill="grey50"),strip.text = element_text(colour = 'white'))




# combine using patchwork with correct annotation.  
fig<-p1/p3/p2+plot_annotation(tag_levels = 'a',tag_prefix = '(', tag_suffix = ')')
fig  

```

